# Data

## Description

Annually, the National Center for Education Statistics (NCES) through its Education Demographic and Geographic Estimate (EDGE) program updates the geographical coordinates (latitude and longitude) of public elementary and secondary schools within the NCES Common Core of Data (CCD). The CCD program gathers administrative and fiscal data from state education agencies about public schools, school districts, and state education offices across the United States. These statistics, sourced from state education agency representatives, encompass essential contact information, directory details for schools and districts, student body composition, instructor numbers, grades taught, and other administrative particulars. The provided physical locations of schools and administrative offices are instrumental in determining the precise point locations for CCD schools and agencies.

Data Link: https://nces.maps.arcgis.com/home/item.html?id=e3cbd4087f144ec5ac3b31094ec84199&sortOrder=desc&sortField=defaultFSOrder#overview

Download Link: https://data-nces.opendata.arcgis.com/datasets/e3cbd4087f144ec5ac3b31094ec84199/explore

Data Format: The dataset is provided in a tabular format.

Dimensions: 100722 rows and 77 features

Field Types: The field types included in the data are String, Integer and latitude and longitude are represented as double.

## Research Plan

We can go into more detail about how each of these factors helps to answer the study questions about educational fairness and geographical differences in US public schools given the dataset columns that are provided:

1. *Geographical Information:*
   - *Location Address, City, State, ZIP Code, Latitude, and Longitude:* It is possible to identify schools in urban, suburban, or rural locations thanks to the extensive geographic information about each school that is provided in these columns. The distribution of schools across the nation can be mapped and visualised through the use of latitude and longitude in spatial analysis.
   - *County Name:* This variable enables more investigation of the geographic setting, taking into account features at the county level that could affect educational results.

2. *Educational Equity:*
   - *Total of Free Lunch and Reduced-Price Lunch Eligible, Free Lunch Program, Reduced-Lunch Program:* These columns shed light on pupils' socioeconomic circumstances. Examining these characteristics can assist in determining the student population's economic background, which is important information for comprehending educational equity.
   - *Title I Eligible, School-wide Title I:* These columns provide information about a school's eligibility for Title I financing, which is frequently linked to educational institutions that serve kids from low-income backgrounds.
   - *Total Students, All Grades:* One key indicator of the overall size of the institution is the total number of students. It is used as the denominator in the computation of several ratios, such as the student-teacher ratio.
   - *Total Teachers, Student-Teacher Ratio:* Details about each school's instructional materials are provided in these columns. In order to ensure educational equity, it is essential to assess the sufficiency of teacher resources, which can be done by analysing these variables.

3. *Racial Disparities:*
   - *Columns for Students of Different Racial Backgrounds (e.g., AM, AS, BL, HP, HI, TR, WH):* A breakdown of student enrollment by gender and race is shown in these columns. It is possible to investigate racial differences in enrollment, performance, and resource access through the analysis of these variables.

4. *School Characteristics:*
   - *Whether a Charter School, Magnet School Indicator, Virtual School Status:* The sort of school is described in these columns. Understanding how various school types contribute to patterns connected to equity and geography can be gained through the analysis of these factors.
   - *School Level:* This variable provides information about whether disparities and equity issues fluctuate at different educational levels by classifying schools into elementary, middle, and high schools.
   - *Locale Code:* This code categorises the school's location, specifying whether it is in a town, city, suburb, or rural area. Investigating the potential effects of the local environment on educational outcomes requires this information.

These factors can be combined and analysed to answer the research questions given in the introduction. Finding trends, relationships, and disparities pertaining to geography, race, and educational equity can be accomplished through the use of statistical modelling, spatial analysis, and data visualisation approaches. A thorough investigation of the intricacies of the US public education system can be supported by the dataset's rich information on student demographics, school features, and geographic location.

## Missing value analysis

In our analysis, we are specifically examining missing values in the "total" column, which signifies the total number of students enrolled in each school. This column holds significant importance because the absence of total student numbers for a school renders the analysis of student figures across different races meaningless. Thus, our initial focus is on analyzing the percentage of missing values for total student counts in each state.

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library(openintro)
library(dplyr)
library(tidyr)
library(tibble)
library(ggplot2)
library(choroplethr)
library(scales)

```


```{r}
data = read.csv("Data/Public_School_Characteristics_2020-21.csv")

missing_state <- data %>% 
    group_by(LSTATE) %>% 
    summarise(sum_na = round(mean(is.na(TOTAL)),4)*100)

missing_state <- missing_state %>% 
                  transmute(region = tolower(usdata::abbr2state(LSTATE)), 
                            value = sum_na)

state_choropleth(missing_state %>% drop_na(),
                 title = "Percentage of missing values for total students by State",
                 legend = "Percentage")
```


The U.S. education system categorizes grades into corresponding school levels, such as Adult Education (AE), Prekindergarten (PK) for Prekindergarten, Prekindergarten (PK) to Grade 5 (G5) for Elementary, Grades 6 to 8 (G6-G8) for Middle School, and Grades 9 to 12 (G9-G12) for High School and Secondary. In our upcoming analysis, we've visualized missing values for each grade based on their associated school levels. To ensure a focused examination, it's crucial to recognize the specific grades within each school level. Notably, a school level may not encompass all grades, as indicated in the graphical representation. For instance, when analyzing the "High School" category, missing values in columns "G09" through "G12" are of interest, recognizing that a high school is not expected to have students enrolled in, for instance, "G01." and hence the missing values in "G01" can be ignored. This nuanced approach enables a more accurate assessment of missing value percentages for relevant grades within a particular school level in the dataset of public schools.

```{r}
data1 = data %>% drop_na(TOTAL)

data1 <- data1 %>% select(23,34:50)

missing_counts <- data1 %>%
  group_by(SCHOOL_LEVEL) %>%
  summarize(across(-starts_with("SCHOOL"), ~mean(is.na(.))*100, .names = "{col}"))

missing_counts_long <- pivot_longer(missing_counts, cols = -starts_with("SCHOOL"), names_to = "Grades", values_to = "MissingCountPercent")

custom_grades <- c("PK","KG","G01","G02","G03","G04","G05","G06","G07","G08","G09","G10","G11","G12","G13","UG","AE")

custom_SCHOOL_LEVEL <- c("Prekindergarten","Elementary","Middle","Secondary","High","Ungraded","Adult Education","Not Applicable","Other")

missing_counts_long$Grades <- factor(missing_counts_long$Grades, levels = custom_grades)
missing_counts_long$SCHOOL_LEVEL <- factor(missing_counts_long$SCHOOL_LEVEL, levels = custom_SCHOOL_LEVEL)

# Create a heatmap
ggplot(missing_counts_long, aes(x = SCHOOL_LEVEL, y = Grades, fill = MissingCountPercent))+
  geom_tile() +
  scale_fill_gradient() +
  labs(title = "Heatmap of Missing Values per Columns per Category",
       x = "School Levels",
       y = "Grades") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90))

```

